### 2020-02-08 thesis/paper.tex
more

\section{Other Methods}
\subsection{Dimensionality Reduction}\label{sec:dim_reduction}
PCA \cite{pca}

 t-SNE \cite{Maaten2008VisualizingDU}

@@

As the data may still contain unwanted dimensions after loading cleanup may need to be done. To reduce the dimensionality the simplest method is to select the ``Dim. Remover'' tool in the top right and define the index of the dimension to be removed. As more sophisticated methods PCA \cite{pca} and t-SNE \cite{Maaten2008VisualizingDU} are also available. PCA calculates the principal components of the data and projects it onto the $k$ strongest ones leaving us with $k$ dimensions. $k$ can easily be defined in the settings of PCA by filling in the dimensions field, as seen in Figure \ref{fig:loaded-data}. t-SNE works similarly to PCA, has more options though and even a parallel implementation. For more information on those methods see Section \ref{sec:dim_reduction}.

@@

@@

\subsection{MDS Plot}
The Multi-Dimensional Scaling (MDS) plot shows points such that their actual distances are approximately represented in a fixed dimensional space. This is done with the equally named method multi-dimensional scaling \cite{mds} for which a maximum number of dimensions $d$ must be defined and the input requires a distance matrix for the points to be layed out. This method is implemented using the Java Smile \cite{javasmile} library and trying to use as many dimensions as possible. With $d$ set, it aims to compute approximate coordinates for the points by trying to keep the distances in a $d$ dimensional space as true to the inserted distance matrix as possible. The resulting coordinates, similarly to PCA, also capture as much information as possible in the leading dimensions, resulting in the first two being the best choice to visualize in 2D. This is the case as MDS actually computes the same result as PCA whenever the distances are Euclidean. To show this information, the created points are loaded into a scatter plot and per default these first two dimensions are shown. Whenever more than two dimensions are available the user can change the shown dimensions by clicking on the corresponding axis and choosing another dimension.

>>>>

\section{Other Methods}

\subsection{Dimensionality Reduction}\label{sec:dim_reduction}

\subsubsection{Principal Component Analysis}
 pca \cite{pca}

\subsubsection{t-Distributed Stochastic Neighbor Embedding}
 t-SNE \cite{Maaten2008VisualizingDU}

\subsubsection{Multi-Dimensional Scaling}
MDS \cite{mds}

@@

As the data may still contain unwanted dimensions after loading cleanup may need to be done. To reduce the dimensionality the simplest method is to select the ``Dim. Remover'' tool in the top right and define the index of the dimension to be removed. As more sophisticated methods Principal Component Analysis (PCA) \cite{pca} and t-Distributed Stochastic Neighbor Embedding (t-SNE) \cite{Maaten2008VisualizingDU} are also available. PCA calculates the principal components of the data and projects it onto the $k$ strongest ones leaving us with $k$ dimensions. $k$ can easily be defined in the settings of PCA by filling in the dimensions field, as seen in Figure \ref{fig:loaded-data}. t-SNE works similarly to PCA, has more options though and even a parallel implementation. For more information on those methods see Section \ref{sec:dim_reduction}.

@@

\begin{figure}[h]
	\centering
	\includegraphics[scale=.45]{data-view}
	\caption{PLACEHOLDER: MDS Plot in Meta-View }
	\label{fig:mdsplot}
\end{figure}

Above the reachability plot in the top right the MDS plot is visible, it can be seen in Figure \ref{fig:mdsplot}. It shows an approximate representation of the clusterings in a finite dimensional space, aiming to stay as true as possible to the distance between them. The computation of this representation uses Multi-Dimensional Scaling (MDS) \cite{mds} which also gives this plot its name. When ground truth is available, it is shown in this plot as a star and the other clusterings are colored according to the threshold in the OPTICS plot. This visualization is useful as the user can, similarly to the heat map, explore which clusterings are close to each other and where the ground truth lies in this representation space. An ideal sampling of the solution space could be claimed if the ground truth is central and as the represented space approximates the solution space, finding the ground truth in the center and surrounded by solutions should be an indicator of good sampling. In the MDS plot the user can also once again select points by clicking, additionally dragging and selecting multiple points at once is possible as well. Besides this, all functionality available in simple scatter plots can also be used here.

@@

\subsection{MDS Plot}
The Multi-Dimensional Scaling (MDS) plot shows points such that their actual distances are approximately represented in a fixed dimensional space. This is done with the equally named method multi-dimensional scaling \cite{mds} for which a maximum number of dimensions $d$ must be defined and the input requires a distance matrix for the points to be layed out. This method is implemented using the Java Smile \cite{javasmile} library and trying to use as many dimensions as possible. With $d$ set, it aims to compute approximate coordinates for the points by trying to keep the distances in a $d$ dimensional space as true to the inserted distance matrix as possible. The resulting coordinates, similarly to PCA, also capture as much information as possible in the leading dimensions, resulting in the first two being the best choice to visualize in 2D. This is the case as MDS actually computes the same result as PCA whenever the distances are Euclidean. To show this information, the created points are loaded into a scatter plot and per default these first two dimensions are shown. Whenever more than two dimensions are available the user can change the shown dimensions by clicking on the corresponding axis and choosing another dimension. The selection of points can be done by either clicking or dragging on the canvas of the scatter plot. When clicking the closes point is selected and when dragging all points within the drawn rectangle are selected.
